---
title: "New Brunswick COVID-19 Rates"
author: "Original work by [Gordon Shotwell](https://shotwell.ca/), Adaptation to New Brunswick data by [Justin d'Entremont](mailto:justin5824@gmail.com)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  html_notebook: 
    highlight: haddock
---

```{r setup, include=FALSE}
library(tidyverse)
library(RcppRoll)
library(plotly)
library(DT)
#Load from manuaaly updated CSV file
nb_daily_covid <- read_csv("nb_daily_covid.csv", col_types = cols(SOE = col_integer()))

covid <- nb_daily_covid %>% 
  mutate(date = lubridate::ymd(date),
         five_day_cases = roll_sum(daily_pos, 5, fill = NA, align = "right"))

covid$no_action <- 1
for (i in 2:nrow(covid)) {
  covid$no_action[i] <- covid$no_action[(i - 1)] * 1.32
}
covid$no_action <- round(covid$no_action, 0)
```
# Data 

The data comes from [daily reports](http://gnb.ca/coronavirus) from the New Brunswick Provincial Government which I have recorded in a [Github Repo](https://github.com/lubindent/nb_covid). *Some data has been taken from media reports to fill in blanks in testing numbers prior to March 28th.*

This page current displays data from `r format(covid$date[min(covid[1])], format="%A, %B %d")` up until `r format(covid$date[max(covid[1])], format="%A, %B %d %Y")`.

---

# Charts

### Cases

The first plot is a plot of total positive tests over time. The red line shows a 32% daily growth rate, which is about what would've happened had we not undertaken any public policy interventions. The black line is the actual cases. 

```{r plots, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}
theme_set(theme_minimal())

covid %>% 
  select(date, actual_cases = positives, no_action) %>% 
  pivot_longer(c("actual_cases", "no_action"), names_to = "type", values_to = "cases") %>% 
  mutate(Scenario = factor(type, levels = c("no_action", "actual_cases"))) %>% 
  ggplot(aes(x  = date, y = cases, group = Scenario, colour = Scenario)) +
  scale_color_manual(values = c("red", "black"), labels = c("No Intervention", "Actual Cases")) +
  geom_line() +
  geom_vline(xintercept = lubridate::ymd("2020-03-19"), 
             colour = "gray") +
  annotate("text", 
           x = lubridate::ymd("2020-03-19") - 0.4, 
           y = 350, 
           label = "State of Emergency Declared", 
           color = "gray",
           angle = 90) +
  labs(y = "Positive Tests",
       x = "Date",
       title = "New Brunswick COVID Cases") 
```

---
```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}
#Calculate hospitalization rates
in_hosp <- covid[max(covid[1]),]$in_hospital
active_case <- covid[max(covid[1]),]$active
current_hosp_rate <- round(in_hosp / active_case * 100,2)

tot_hosp <- covid[max(covid[1]),]$total_hospitalized
tot_case <- covid[max(covid[1]),]$positives
tot_hosp_rate <- round(tot_hosp / tot_case * 100,2)

```


It's also useful to track active cases, cases who have recovered or are currently in hospital. 

Take note that the current hospitalization rate $\frac{\text{# in hospital}}{\text{# active cases}}$ is `r current_hosp_rate`%  and the total hospitalization rate $\frac{\text{# total admited to hospital}}{\text{# total cases}}$ is `r tot_hosp_rate`%.

```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}

covid %>% 
  select(date, total = positives, recovered, in_hospital) %>% 
  mutate(active = total - recovered) %>% 
  #select(-total) %>% 
  pivot_longer(-date, names_to = "group", values_to = "number") %>% 
  mutate(group = factor(group, levels = c("total", "active",  "recovered", "in_hospital"),
                        labels = c("Total Cases",  "Active Infections",  "Recovered", "Currently In Hospital"))) %>% 
  ggplot(aes(x = date, y = number, group = group, colour = group)) +
  geom_line() +
  labs(y = "Number of people",
       x = "Date",
       title = "New Brunswick COVID Cases")

```

---

### Growth


I stole this plot from an excellent [data visualization](https://aatishb.com/covidtrends/) of COVID trends. Each point is a day and it shows the sum of new cases in the previous five days against the number of total cases. I chose a five day sum because that's about the mean incubation period for COVID-19.

I encourage anyone to watch **[this YouTube video](https://www.youtube.com/watch?v=54XLXg4fYsc)** to get a better understanding how to read this chart. In brief:

- When the line is flat, infections are growing at a steady (but not exponential) rate. 
- When the line is sloped up, the number of cases are growing exponentialy.
- When the line is sloped down, rate of new of cases appearing is decreasing.
- When the line is headed straight down, we will seem to have a handle on the spread of diease and are on the way down the "curve".

```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}
log_graph <- ggplot(covid, aes(x = positives, y = five_day_cases)) +
  geom_point() +
  geom_line() +
  labs(y = "Cases in Previous Five Days",
       x = "Total Cases",
       title = "New Brunswick COVID cases")
log_graph + scale_y_log10()
```

---

### Number of Tests

Next we look at tests. The first one shows how there is a lot of missing (unpublished) data in what is curently published on the GNB website, particurarly for early on in the outbreak. GNB only started publishing daily testing numbers beginning on March 28th.

```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}

covid %>% 
  select(date, total_tests) %>% 
  pivot_longer(-date, names_to = "group", values_to = "number") %>% 
  mutate(group = factor(group, levels = c("total_tests"),
                        labels = c("Total Tests Conducted"))) %>% 
  ggplot(aes(x = date, y = number, group = group, colour = group)) +
  geom_line() +
  labs(y = "Number of Tests",
       x = "Date",
       title = "New Brunswick - Total Number of COVID Tests Conducted")

```



```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}
ggplot(covid, aes(x  = date, y = daily_tests)) +
  geom_point() +
  geom_line() +
  labs(y = "Tests per day",
       x = "Date",
       title = "New Brunswick - Number of COVID Tests Conducted Per Day") +
  ylim(0, NA)

```

---

The last thing to look at is the positive rate for our testing. This should be read in conjunction with the previous chart because if we test a lot of people, the positive rate will probably go down even though nothing changed about the disease. The thing to pay attention here is if these two move in the same direction, if the number of tests goes up and the positive rate also goes up that's bad, if they both go down that's good.

```{r, echo = FALSE, warning=FALSE, message=FALSE, dev='svg'}
ggplot(covid, aes(x  = date, 
                  y = daily_pos / daily_tests,
                  size = daily_tests)) +
  geom_point() +
  scale_y_continuous(labels = scales::percent)+
  labs(y = "Positive rate",
       x = "Date",
       title = "New Brunswick COVID tests") 
```

---

# Raw Data

Here is the raw data collected, please let me know if I've made any mistakes in recording this data (Justin5824@gmail.com). 

```{r, layout="l-body-outset", echo = FALSE}
covid_df <- as.data.frame(covid)
library(rmarkdown)
paged_table(covid)
```

